---
title: "High Dimensional Data Analysis - Group Assignment 18"
author: "Group 18"
date: "08/09/2021"
output:
  bookdown::html_document2:
    fig_height: 5
    fig_width: 8
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    code_folding: show
    theme: readable
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
library(tidyverse)
library(dplyr)
library(visdat)
library(splitstackshape)
library(lubridate)
library(patchwork)
bankruptcy <- read_rds(here::here("data/Bankruptcy.rds"))
```


# Preliminary Data Analysis


Before carrying out further analysis of the data, let us conduct some preliminary data analysis. The function skim() from the skimr R package gives an overview of the data. From the summary shown below, we can see that the data is a high dimensional dataset with 21 variables, out of which 6 are character variables and 15 are numeric variables. It can be observed that even though the complete rate of  variable `FirmEnd` is 100%, there are quite a number of empty values present which are essentially `NULL` values. Therefore, the first step of our data preparation would be to convert these empty values to missing `NA` values. 

```{r prelim-skimr}
skimr::skim(bankruptcy)
```

```{r prelim-firmend}
bankruptcy <- bankruptcy %>%
  mutate(FirmEnd = ifelse(FirmEnd == "", 
                          NA, 
                          FirmEnd))
```

Before we delve in deeper, the data credibility issues are checked by confirming if the `DaysIn`, `EmplUnion`, `Employees`, `HeadCourtCityToDE`,`MonthFiled`, `YearFiled` and `HeadCityPop` are non-negative values. We have also found that there are observations where the `EmplUnion` values are more than `Employees` which was removed from the data and that certain companies have 1 Employee and 1 EmplUnion values as shown below, which is suspicious but since there is not any concrete evidence that these observations pose data credibility issues, these observations were not excluded for the analysis. 

```{r}
bankruptcy%>%
  filter(EmplUnion>=Employees)%>%
  select(Name,Employees,EmplUnion)

bankruptcy <-  bankruptcy %>%
filter(Name != "Promus Companies Inc. (Harrahs Jazz Co. only)")
```


Looking at the data again through the glimpse() function from dplyr package, we noticed that `SICMajGroup`,the industry classification code,has a lengthy name, therefore inorder to keep it less complicated and more identifyable, we have separated it to the code into a new factor variable `SIC` and retained its meaning in the `SICMajGroup`. 

```{r prelim-sic}
glimpse(bankruptcy)
bankruptcy <- bankruptcy %>%
  separate(SICMajGroup, 
           into = c("SIC", "SICMajGroup"), 
           sep = "\\s", 
           extra = "merge") %>%
  mutate(SIC = as.factor(SIC))
```


Looking into the data summary again, we can see that some of the numeric variables in the data such as `DaysIn`, `Ebit`, `Employees`,`EmplUnion`, `HeadCourtCityToDE`, `Liab` and `Sales` have numerous missing values that is 6.6% of the data is missing. The missing values in the data has been visualized as shown in \@ref(fig:vismiss). Throughout our strategy, we have tried to retain the data as much as possible while maintaining high data quality and credibility. 

It can be observed that `FirmEnd` has the highest number of missing values with 94.25% of the values as missing followed by `EmplUnion` with 31.03%. In order to deal with these missing values, the strategy employeed is to remove the variables `FirmEnd` and `EmplUnion`. As the variable `FirmEnd` depicts a short description of the event that ended the firm’s existence, it does not provide significant value to the analysis. Therefore it can be excluded from further analysis. Similarly `EmplUnion` is removed due to the fact that `Employees` and `EmplUnion` are closely related that is `EmplUnion` will always be a subset of `Employees`, therefore removing `EmplUnion` which has too many missing would not affect our analysis significantly as the variable Employees explains the aspect of the number of employees present in the organisation before bankruptcy. 

```{r prelim-vismiss}
vis_miss(bankruptcy)+
  ggtitle("Missing values in the Data")


bankruptcy<- bankruptcy%>%
  select(-FirmEnd,-EmplUnion)
```

Moving on to the other variables, we have employeed the strategies listed below, to deal with the missing values in each variable.

-  **DaysIn**

Exploring the `DaysIn` variable further, it was observed that there are 4 companies whose values for `DaysIn` as missing. Based on the publicaly available data,

- [AP Industries, Inc.](https://casetext.com/case/in-re-ap-industries-inc?__cf_chl_jschl_tk__=pmd_0e9LcvX3fGC0nT5WJVBShbHDvK9GY2AMZnekCjslbN0-1631333530-0-gqNtZGzNAjujcnBszQd9) filed the bankruptcy on 27/03/1990, disposition made in 26/07/1990, therefore `DaysIn` can be encoded equivalent to 121 days

- [Daisy Systems Corp.](https://caselaw.findlaw.com/us-9th-circuit/1020075.html) filed the bankruptcy on 30/05/1991, the court decided on 24/09/1996 , therefore `DaysIn` can be encoded equivalent to 1944 days. 

However, the data for **Hunt International Resources Corp.** and **McCrory Corp.** was not available, therefore we have imputed the days in bankruptcy, based on median of the length of bankruptcy process in the industry classification they below to (Hunt International Resources Corp., SIC = 20; McCrory Corp., SIC = 53). 



```{r prelim-daysin}
na_daysin <- filter(bankruptcy, is.na(DaysIn))
na_daysin%>%
  select(c(Name,DaysIn))
```

```{r}
median_daysin<- bankruptcy %>% group_by(SIC) %>%
  summarise(median = median(DaysIn, na.rm = TRUE)) %>% 
    dplyr::filter(SIC %in% c(20, 53))
```


```{r prelim-daysinimpute}
bankruptcy <- bankruptcy %>%
  mutate(DaysIn = ifelse(Name == "Hunt International Resources Corp.", 305,
                                    ifelse(Name == "AP Industries, Inc.", 121,
                                           ifelse(Name == "Daisy Systems Corp.", 1944,
                                                  ifelse(Name == "McCrory Corp.", 683, DaysIn)))))

summary(bankruptcy$DaysIn)

bankruptcy%>%
  ggplot(aes(x=DaysIn))+
  geom_histogram(fill="blue",alpha=0.7)+
  theme_minimal()+
  ggtitle("Histogram of DaysIn")+
  ylab("Count")

```

The summary statistics and the histogram of the `DaysIn` variable after imputation, suggests no suspicious outliers or anomalies as the bankruptcy has be a lengthy ordeal. 


- **HeadCourtCityToDE** 

- The missing values in the variable shown in the table below, are imputed using the values in `CityFiled`, `DENYOther`, or `HeadStAtFiling`. Considering the data on city from the company address available publicaly and the `CityFiled`, the distances between these cities were found and imputed into the data accordingly.  This brings us to the conclusion that the `HeadCourtCityToDE` for [Divi Hotels, N.V.](https://www.bloomberg.com/profile/company/DVH:US) is 1126 miles where as for [Loewen Group, Inc](https://www.sedar.com/DisplayProfile.do?lang=EN&issuerType=03&issuerNo=00001601) (British Columbia to Wilmington)  and [Philip Services Corp.](https://www.sedar.com/DisplayProfile.do?lang=EN&issuerType=03&issuerNo=00001613) (Ontario to Wilmington) is 2942 and 1234 miles respectively. 


```{r prelim-headquart}
na_distance <- bankruptcy %>%
  filter(is.na(HeadCourtCityToDE))

na_distance%>%
  select(Name,HeadCourtCityToDE,CityFiled,DENYOther,HeadStAtFiling)
```

```{r}
bankruptcy <- bankruptcy %>%
  mutate(HeadCourtCityToDE = ifelse(Name == "Divi Hotels, N.V.", 1126,
                                    ifelse(Name == "Loewen Group, Inc.", 2942,
                                           ifelse(Name == "Philip Services Corp. (1999)", 1234,
                                                  HeadCourtCityToDE))))

summary(bankruptcy$HeadCourtCityToDE)

bankruptcy%>%
  filter(HeadCourtCityToDE==1)%>%
  select(Name,HeadCourtCityToDE,CityFiled,DENYOther,HeadStAtFiling)
```

```{r prelim-headquarthisto}
bankruptcy%>%
  ggplot(aes(x=HeadCourtCityToDE))+
  geom_histogram(fill="blue",alpha=0.7)+
  theme_minimal()+
  ggtitle("HeadCourtCityToDE")+
  ylab("Count")
```
Exploring the summary statistics, it was observed that the minimum distance is 1 , when inspected the state of headquarter and the city filed its in the same state therefore it does not pose a data credibility issue. From the \@ref(fig:headquarthisto), we can observe that there are no severe outliers present. 

- **Sales and Employees**


    
With regards to the `Employees` variable, it was observed that there is a single observation that is missing data on its employees. Under closer examination of the `Sales` Variable, we observed that the only firm that was missing data on its `Sales` was same firm that has missing values in Employees Variable. On closer inspection of this firm, we have found that the firm has missing values on the variable `Ebit` as well, therefore considering the fact that there is only a single observation that is missing in `Employees` and `Sales`, along with the fact that these observations are the same and the observation also contains missing values for `Ebit`, we are removing the observation from the data for further analysis. 
```{r}

na_sales <- filter(bankruptcy, is.na(Sales))
na_sales%>%
  select(Name,Sales)
```

```{r}
na_employees <- filter(bankruptcy, is.na(Employees))

na_employees%>%
  select(Name,Employees)

```
    
```{r}

options(scipen = 999)
bankruptcy%>%
  filter(Name=="County Seat, Inc.")%>%
  select(Name,Sales,Employees,Ebit)

bankruptcy<- bankruptcy%>%
  filter(!is.na(Sales))


```


**Ebit and Liab**


The missing values in `Liab` and `Ebit` was treated by dropping the missing observations, as the missing values in each of the variables were below 10% and out of the 39 rows where either one of the two variables were missing, 8 of the observations have missing values on both `Liab` and `Ebit`.

```{r}
all_liab_ebit_missing<- bankruptcy %>%
  filter(is.na(Ebit) |is.na(Liab))

bank_data_op1 <- bankruptcy %>%
  filter(!is.na(Ebit) & !is.na(Liab))
data_label <- bank_data_op1%>%
  filter(Ebit==max(Ebit)|Liab==max(Liab)|Sales==max(Sales))%>%
  select(Name,Ebit,Liab,Sales)


```


It was also noticed that variables such as that `DENYOther`, `MonthFiled` and `YearFiled` ought to be factor as mentioned in the data description are depicted as numeric. These variables are then converted to factor variables as shown in figure \@ref(fig:cleandata)

```{r cleandata}

data_clean <- bank_data_op1 %>%
  mutate(DENYOther = as.factor(DENYOther),
         MonthFiled = as.factor(MonthFiled),
         YearFiled = as.factor(YearFiled))
row.names(data_clean) <- data_clean$Name
vis_dat(data_clean)
```

The data was then checked for outliers, even though we havent found suspicious outliers in majority of the variables, we have however found outliers in `Ebit`, `Liab` and `Sales`. Interestingly, these values belong to a single firm called **Texaco Inc.** This will be discussed further in the sections below.


```{r}
p3 <- data_clean%>%
  ggplot()+
  geom_histogram(aes(x=Sales),fill="blue",alpha=0.7)+
    geom_text(data=data_label,aes(x=Sales,y=20,label=Name))+
  theme_minimal()+
  ggtitle("Sales")+
  ylab("Count")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p1<- data_clean%>%
  ggplot()+
  geom_histogram(aes(x=Ebit),fill="blue",alpha=0.7)+
  geom_text(data=data_label,aes(x=Ebit,y=20,label=Name))+
  theme_minimal()+
  ggtitle("Ebit")+
  ylab("Count")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

p2 <- data_clean%>%
  ggplot()+
  geom_histogram(aes(x=Liab),fill="blue",alpha=0.7)+
  geom_text(data=data_label,aes(x=Liab,y=20,label=Name))+
  theme_minimal()+
  ggtitle("Liab")+
  ylab("Count")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```



```{r}
p1+p2+p3

```

In order to gain insights from the data, we have further explored the data 